OSM-GB - basic analysis of the dataset
========================================================

The file "great-britain-latest.osm.pbf" was downloaded from [geofabrik](http://download.geofabrik.de/europe/great-britain.html) on 10th March 2014.

This file has a compressed size of 611 Mb, and is constantly growing due to new 
additions to the map from the OSM community. Clearly, this is a large file to manage, 
so the first stage was to filter out all of the information except for cycleways. 
This was done using the following command.

```{bash}
osmosis --read-pbf great-britain-latest-10-03-2014.osm.pbf --tf accept-ways highway=cycleway --used-node --write-xml gb-cways.osm
```

This command took 5 minutes to run on a 3rd generation Core i5 laptop computer with 12 Mb of RAM.
The resulting file was 260 Mb in size. It was decided that a further subset would be taken of this file, to facilitate fast preliminary analysis. To start with, we took the bounding box 
encapsulating Leeds and Nottingham, the home locations of the authors and where
the cycle network is best understood in intuitive terms.

```{bash}
osmosis --read-xml gb-cways.osm --bounding-box top=53.2 left=-1.8 bottom=52.7 right=-1.3 --write-xml NottLeedsCways.osm
```
This resulted in a file that was a more manageable size: 4 Mb

## Analysis in R

To read in the file, the following command was run, having first loaded the 
osmar package.

```{r}
# cway <- get_osm(complete_file(), source = osmsource_file("/media/SAMSUNG/geodata/osm-data/NottLeedsCways.osm"))
# cwaySp <- as_sp(cway, "lines")
# save.image()
load(".RData")

nrow(cwaySp)
plot(cwaySp)
head(cwaySp@data)

# Let's see the distribution of times of contribution:
hist(cway$nodes$attrs$timestamp, breaks = "month")
```



You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
library(rgeos)
plot(gLength(cwaySp, byid=T), cwaySp$timestamp, xlim=c(0,0.1))
```

