OSM-GB - basic analysis of the dataset
========================================================

This vignette describes how to load-in and undertake basic analysis of cycle 
path data in OSM.

## Accessing the data

### Downloading sections of planet.osm

Planet.osm is the name of the entire world, as represented in a single, large (~32 Gb)
xml file. Clearly this is not easy to download or process so we downloaded sections.

The file "great-britain-latest.osm.pbf" was downloaded from [geofabrik](http://download.geofabrik.de/europe/great-britain.html) on 10th March 2014. Another possible solution would be 
to use the "overpass" API, but this approach has not yet been attempted.

This file has a compressed size of 611 Mb, and is constantly growing due to new 
additions to the map from the OSM community. Clearly, this is a large file to manage, 
so the first stage was to filter out all of the information except for transport routes.



Next, a filter was used to try to extract all the 'cycleways'.
This was done using the following command (note that this is not guaranteed to extract all cycle ways).

```{bash}
osmosis --read-pbf great-britain-latest-10-03-2014.osm.pbf --tf accept-ways highway=cycleway --used-node --write-xml gb-cways.osm
```

This command took around 5 minutes to run on a 3rd generation Core i5 laptop computer with 12 Mb of RAM.
The resulting file was 260 Mb in size. It was decided that a further subset would be taken of this file, to facilitate fast preliminary analysis. To start with, we took the bounding box 
encapsulating Leeds and Nottingham, the home locations of the authors and where
the cycle network is best understood in intuitive terms.

```{bash}
osmosis --read-xml gb-cways.osm --bounding-box top=53.2 left=-1.8 bottom=52.7 right=-1.3 --write-xml NottLeedsCways.osm
```
This resulted in a file that was a more manageable size: 4 Mb.

### Using the Overpass API

The [Overpass API](http://wiki.openstreetmap.org/wiki/Overpass_API) allows specific
parts of the planet.osm file to be accessed and downloaded directly, using its 
own language of filters. To download all cycle paths in the UK, for example, 
the following query can be used:

```{bash}
http://overpass-api.de/api/interpreter?[timeout:600];data=(way(poly:"51.33770 
1.983960 50.10375 0.05807600 49.89032 -1.246113 49.77612 -2.027284 
49.76352 -2.020900 49.76171 -2.031990 49.10263 -1.808598 49.00443 
-1.835368 49.15592 -3.940324 49.70097 -7.024780 52.78017 -5.441616 
53.77268 -5.206178 54.86510 -5.488813 55.33747 -6.208707 55.45018 
-6.730561 57.73276 -9.408655 58.20977 -9.077830 60.34515 -4.158581 
61.13564 -1.637515 61.1030 -0.2670265 59.76591 0.1278457 59.42198 
0.09781600 55.79959 0.7998970 54.43427 1.702395 52.58000 
2.250000")[highway=cycleway];);<;>;out meta;
```

Many thanks to Toni Hernandez at the 
[SIGTE](http://www.sigte.udg.edu/sigte_en/) service of the University 
of Girona for demonstrating this method. 

This resulted in a 16 Mb .pbf file. This was converted to an osm file using osmconvert:

```{r}
osmconvert GB_cycleways.pbf >GB_cycleways.osm
```

The resulting file was 230 Mb, similar in size as that obtained from the 
osmosis filter method, but without the superfluous points. 

### Accessing the data using WFS in QGIS

WFS is the Web Feature Service, a way to serve geographic data
over the internet. OSMGB has its very own server based in Nottingham that 
can serve all your data needs, as described in an excellent 
[article](http://elogeo.nottingham.ac.uk/xmlui/bitstream/url/157/1/FOSS4Gworkshop.pdf) 
on the 
matter. 

### Relations

In fact, not all cycle paths are marked with the tag highway='cycleway', 
let alone bicycle='yes'. Instead, 
[the OSM community](http://wiki.openstreetmap.org/wiki/Cycle_routes) 
prefers users to mark cycle paths with relations.


### Accessing the data in PostGIS

For efficient and stable processing with a a wide range of geo-functionallity 
the spatial enhancement of Postgres, PostGIS can be used. The raw data can be 
read into a PostGIS database as follows:

```{bash}
createdb Cyway # create the database called osmTutdb

psql osmTutdb -c "create extension postgis" # add PostGIS functionality

psql osmTutdb -c "create extension pgrouting" # add pgrouting functionality (not essential)

osm2pgsql --extra-attributes -S default.style -d Cycway GB_cycleways.pbf 
```

In the above block of code, the final command is the most important, 
ensuring that the timestamp and user id of the bicycle paths is included.

## Analysis in R

To read in the file, the following command was run, having first loaded the 
osmar package.

```{r}
# cway <- get_osm(complete_file(), source = osmsource_file("/media/SAMSUNG/geodata/osm-data/NottLeedsCways.osm"))
# cwaySp <- as_sp(cway, "lines")
# save.image()
load(".RData")

nrow(cwaySp)
plot(cwaySp)
head(cwaySp@data)

# Let's see the distribution of times of contribution:
hist(cway$ways$attrs$timestamp, breaks = "month")
```

g

You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
library(rgeos)
plot(gLength(cwaySp, byid=T), cwaySp$timestamp, xlim=c(0,0.1))
```

Only layers in current visible 

