Classification of cycle network in Open Street Map
========================================================

There are many ways to classify bicycle paths in OSM.

Probably the best public description of the different tags of route
data in OSM and their cyclability is available from
CycleStreets.net: 
[http://www.cyclestreets.net/journey/help/osmconversion/]{http://www.cyclestreets.net/journey/help/osmconversion/}

This demonstrates that the `highway=cycleway` tag is common
(with 460,000 elements) but by no means the only way
of defining bicycle paths. In fact this classification contains 
less than half of all routes designed to be used by cyclists.
More common is the `cycleway=track` tag, with over 500,000
elements belonging to this category.

There are many additional tags which
must be taken into account to load all cycle
paths in a given area have been captured. The
problem becomes more complex when we remember that
the recommended way of classifying cycle paths is via relations
(see ...). Thus, on-road cycle paths may not be visible from
tag data at all - a potentially confusing feature of OSM data
for new users.
With many years of practical
experience in this realm, CycleStreets.net and OpenCycleMap have
devel

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
plot(cars)
```

