Cycling uptake across the UK
========================================================

```{r}
x <- c("rgdal", "dplyr", "plyr", "ggplot2", "rgeos")
lapply(x, require, character.only=T)
la01 <- readOGR("updata/", "ua2001")
length(la01)
head(la01@data)

names(la01)[9:20] <- c("All", "Mfh", "Tram", "Train", "Bus", "Moto", "Car.d", "Car.p", "Taxi", "Cycle", "Walk", "Other")
la01$pCycle <- la01$Cycle / la01$All 
la01@data[la01$pCycle > 0.1,] # do these seem like cycling towns? - YES!

la11 <- readOGR("updata/", "ua2011")
head(la11@data) # no data for Scotland
length(la11) # 326 las in England

354 - 38 + 8

la11.data <- read.csv("updata/all-modes-la-Data_AGE_MTTWRK_UNIT.csv")
head(la11.data[1:6])
la11.data <- la11.data[!la11.data$F560 == "",] # remove na values
la11.data <-la11.data[!grepl("W", la11.data$GEO_CODE),] # remove welsh areas
la11.names <- la11.data[1,]
la11.names[6:18]
names(la11.data)[c(2, 6:18)] <- c("CODE", "All", "Mfh", "Tram", "Train", "Bus", "Taxi", "Moto", "Car.d", "Car.p", "Cycle", "Walk", "Other", "Unemp")
la11.data <- la11.data[-1,] # remove names row
head(la11.data)

length(which(la11.data$CODE %in% la11$CODE))
la11.data[!(la11.data$CODE %in% la11$CODE), 1:10]
la11@data[!(la11@data$CODE %in% la11.data$CODE),]

# substituting cornwall's code
la11@data[grepl("Corn", la11$NAME), ]
la11.data[grepl("Corn", la11.data$GEO_LABEL), ] # that's cornwall
plot(la11[grepl("Corn", la11$NAME), ]) # it's definitely cornwall!
la11.data$CODE <- as.character(la11.data$CODE)
la11.data$CODE[grepl("Corn", la11.data$GEO_LABEL)] <- "E06000052"

# Westminster
la11@data[grepl("Westm", la11$NAME), ]
la11.data[grepl("Westm", la11.data$GEO_LABEL), ] # different code
la11.data$CODE[grepl("Westm", la11.data$GEO_LABEL)] <- "E09000033"

# City of London
la11@data[grepl("London", la11$NAME), ]
la11.data[grepl("London", la11.data$GEO_LABEL), ] # different code
la11.data$CODE[grepl("City of L", la11.data$GEO_LABEL)] <- "E09000033"

la11@data[!(la11$CODE %in% la11.data$CODE),] # check shapes not in data
la11 <- la11[-which(grepl("City of L|Scill", la11$NAME)),]
plot(la11)

head(la11.data)
head(la11@data)
la11@data <- join(la11@data, la11.data)
head(la11@data)

# convert to numbers
la11@data[!grepl("[1-9]", la11$Train),] 
head(la11@data)
for(i in 8:20){
  la11@data[,i] <- as.numeric(as.character(la11@data[,i]))
}
```

We have now loaded the data at the LA level for 2001 and 2011 modal split data.
Are the proportions generated realistic?

```{r}
names(la11)
sum(la11@data[8]) - sum(la11@data[9:19])
sum(la11$Unemp)
sum(la11$All) - sum(la11$Unemp)
sum(la01$All) # this shows the "All" category includes unemployed. Rm them!
la11$All <- la11$All - la11$Unemp
sum(la11$All) / sum(la01$All) # 12% growth - sounds about right

la11$pCycle <- la11$Cycle / la11$All 
la11@data[la11$pCycle > 0.1,] # do these seem like cycling towns? - YES!
```

Let's look at the cross-comparison between LAs in 2001 and 2011.

```{r}
la01@data[! la01$ZONE_LABEL %in% la11$NAME , 1:5 ] # now switch to qgis to merge these
plot(la01[! la01$ZONE_LABEL %in% la11$NAME ,]) # all zones which have changed
# writeOGR(la01[! la01$ZONE_LABEL %in% la11$NAME ,], dsn="updata/", "las01-not-11", "ESRI Shapefile")
la11merge <- merge(la01@data, la11@data, by.x = "ZONE_LABEL", by.y = "NAME", all.x=F)
plot(la11merge$All.x, la11merge$All.y) # this suggests we can just rm some of the zones woop
la11merge[which(la11merge$All.x / la11merge$All.y < 0.75), 1:3]
```

Merge the data from the 2001 census to fit into 2011 las.
First up: Cornwall.

```{r}
la01.pnts <- SpatialPointsDataFrame(coordinates(la01), data=la01@data)
plot(la11[ grepl("Cornw", la11$NAME), ])
proj4string(la01.pnts) <- proj4string(la11)
cornwal01 <- la01.pnts[la11[ grep("Cornw|Sci", la11$NAME), ],]
cornwal01
points(cornwal01)
corRegionel <- c(row.names(cornwal01), "96") # include isle of scilly

la01.pnts@data <- la01@data[c(2,9:23)]
class(la01.pnts@data$Mfh)
colSums(la01.pnts@data[corRegionel,2:16])
la01.pnts$ZONE_LABEL <- as.character(la01.pnts$ZONE_LABEL)
# la01.pnts@data <- rbind(la01.pnts@data, c("Cornwall,Isles of Scilly", colSums(la01.pnts@data[corRegionel, 2:16])))
newCornw <- SpatialPointsDataFrame(coordinates(la11[ grepl("Cornw", la11$NAME), ]), match.ID= F,
                                         data = data.frame(t(c(ZONE_LABEL =1000, colSums(la01.pnts@data[corRegionel, 2:16])))) )
proj4string(newCornw) <- proj4string(la01.pnts)

la01.pnts <- rbind.SpatialPointsDataFrame(la01.pnts, newCornw )

class(la01.pnts@data$Mfh)
la01.pnts@data[nrow(la01.pnts@data),1] <- "Cornwall"

la01.pnts <- la01.pnts[- which(row.names(la01.pnts) %in% corRegionel) ,]
la11@data[ grep("Cornw", la11$NAME), ]
la01.pnts@data[ grep("Cornw", la01.pnts@data$ZONE_LABEL),]
```

Merge the next zone: Wiltshire.

```{r}
plot(la11)
plot(la11[ grep("Wilt", la11$NAME), ], col = "red", add = T)
Wilt <- la01.pnts[la11[ grep("Wilt", la11$NAME), ],]
points(Wilt)
Wilt # names make sense?
Wilt <- row.names(Wilt)
la01.pnts[Wilt,]

newWilt <- SpatialPointsDataFrame(coordinates(la11[ grepl("Wilt", la11$NAME), ]), match.ID= F,
                                         data = data.frame(t(c(ZONE_LABEL =1000, colSums(la01.pnts@data[Wilt, 2:16])))) )
proj4string(newWilt) <- proj4string(la01.pnts)

la01.pnts <- rbind.SpatialPointsDataFrame(la01.pnts, newWilt )
la01.pnts@data[nrow(la01.pnts@data),1] <- as.character(la11@data$NAME[ grep("Wilt", la11$NAME)])

nrow(la01.pnts)
la01.pnts <- la01.pnts[- which(row.names(la01.pnts) %in% Wilt) ,]
nrow(la01.pnts)
la11@data[ grep("Wilt", la11$NAME), ]
la01.pnts@data[ grep("Wilt", la01.pnts@data$ZONE_LABEL),]
points(la01.pnts, col = "green")
```

Merge the next zone: Northumberland

```{r}
plot(la11)
plot(la11[ grep("Northumb", la11$NAME), ], col = "red", add = T)
Northumb <- la01.pnts[la11[ grep("Northumb", la11$NAME), ],]
points(Northumb)
Northumb # names make sense?
Northumb <- row.names(Northumb)
la01.pnts[Northumb,]

newNorthumb <- SpatialPointsDataFrame(coordinates(la11[ grepl("Northumb", la11$NAME), ]), match.ID= F,
                                         data = data.frame(t(c(ZONE_LABEL =1000, colSums(la01.pnts@data[Northumb, 2:16])))) )
proj4string(newNorthumb) <- proj4string(la01.pnts)

la01.pnts <- rbind.SpatialPointsDataFrame(la01.pnts, newNorthumb )
la01.pnts@data[nrow(la01.pnts@data),1] <- as.character(la11@data$NAME[ grep("Northumb", la11$NAME)])

nrow(la01.pnts)
la01.pnts <- la01.pnts[- which(row.names(la01.pnts) %in% Northumb) ,]
nrow(la01.pnts)
la11@data[ grep("Northumb", la11$NAME), ]
la01.pnts@data[ grep("Northumb", la01.pnts@data$ZONE_LABEL),]
points(la01.pnts, col = "green")
```

Durham:

```{r}
plot(la11)
plot(la11[ grep("Durh", la11$NAME), ], col = "red", add = T)
Durh <- la01.pnts[la11[ grep("Durh", la11$NAME), ],]
points(Durh)
Durh # names make sense?
Durh <- row.names(Durh)
la01.pnts[Durh,]

newDurh <- SpatialPointsDataFrame(coordinates(la11[ grepl("Durh", la11$NAME), ]), match.ID= F,
                                         data = data.frame(t(c(ZONE_LABEL =1000, colSums(la01.pnts@data[Durh, 2:16])))) )
proj4string(newDurh) <- proj4string(la01.pnts)

la01.pnts <- rbind.SpatialPointsDataFrame(la01.pnts, newDurh )
la01.pnts@data[nrow(la01.pnts@data),1] <- as.character(la11@data$NAME[ grep("Durh", la11$NAME)])

nrow(la01.pnts)
la01.pnts <- la01.pnts[- which(row.names(la01.pnts) %in% Durh) ,]
nrow(la01.pnts)
la11@data[ grep("Durh", la11$NAME), ]
la01.pnts@data[ grep("Durh", la01.pnts@data$ZONE_LABEL),]
points(la01.pnts, col = "green")
```

Cheshire E:

```{r}
plot(la11)
plot(la11[ grep("Cheshire E", la11$NAME), ], col = "red", add = T)
Chest <- la01.pnts[la11[ grep("Cheshire E", la11$NAME), ],]
points(Chest)
Chest # names make sense?
Chest <- row.names(Chest)
la01.pnts[Chest,]

newChest <- SpatialPointsDataFrame(coordinates(la11[ grepl("Cheshire E", la11$NAME), ]), match.ID= F,
                                         data = data.frame(t(c(ZONE_LABEL =1000, colSums(la01.pnts@data[Chest, 2:16])))) )
proj4string(newChest) <- proj4string(la01.pnts)

la01.pnts <- rbind.SpatialPointsDataFrame(la01.pnts, newChest )
la01.pnts@data[nrow(la01.pnts@data),1] <- as.character(la11@data$NAME[ grep("Cheshire E", la11$NAME)])

nrow(la01.pnts)
la01.pnts <- la01.pnts[- which(row.names(la01.pnts) %in% Chest) ,]
nrow(la01.pnts)
la11@data[ grep("Cheshire E", la11$NAME), ]
la01.pnts@data[ grep("Cheshire E", la01.pnts@data$ZONE_LABEL),]
points(la01.pnts, col = "green")
```

Cheshire W:

```{r}
plot(la11)
plot(la11[ grep("Cheshire W", la11$NAME), ], col = "red", add = T)
Chest <- la01.pnts[la11[ grep("Cheshire W", la11$NAME), ],]
points(Chest)
Chest # names make sense?
Chest <- row.names(Chest)
la01.pnts[Chest,]

newChest <- SpatialPointsDataFrame(coordinates(la11[ grepl("Cheshire W", la11$NAME), ]), match.ID= F,
                                         data = data.frame(t(c(ZONE_LABEL =1000, colSums(la01.pnts@data[Chest, 2:16])))) )
proj4string(newChest) <- proj4string(la01.pnts)

la01.pnts <- rbind.SpatialPointsDataFrame(la01.pnts, newChest )
la01.pnts@data[nrow(la01.pnts@data),1] <- as.character(la11@data$NAME[ grep("Cheshire W", la11$NAME)])

nrow(la01.pnts)
la01.pnts <- la01.pnts[- which(row.names(la01.pnts) %in% Chest) ,]
nrow(la01.pnts)
la11@data[ grep("Cheshire W", la11$NAME), ]
la01.pnts@data[ grep("Cheshire W", la01.pnts@data$ZONE_LABEL),]
points(la01.pnts, col = "green")
```

Shropshire:

```{r}
plot(la11)
plot(la11[ grep("Shrop", la11$NAME), ], col = "red", add = T)
Chest <- la01.pnts[la11[ grep("Shrop", la11$NAME), ],]
points(Chest)
Chest # names make sense?
Chest <- row.names(Chest)
la01.pnts[Chest,]

newChest <- SpatialPointsDataFrame(coordinates(la11[ grepl("Shrop", la11$NAME), ]), match.ID= F,
                                         data = data.frame(t(c(ZONE_LABEL =1000, colSums(la01.pnts@data[Chest, 2:16])))) )
proj4string(newChest) <- proj4string(la01.pnts)

la01.pnts <- rbind.SpatialPointsDataFrame(la01.pnts, newChest )
la01.pnts@data[nrow(la01.pnts@data),1] <- as.character(la11@data$NAME[ grep("Shrop", la11$NAME)])

nrow(la01.pnts)
la01.pnts <- la01.pnts[- which(row.names(la01.pnts) %in% Chest) ,]
nrow(la01.pnts)
la11@data[ grep("Shrop", la11$NAME), ]
la01.pnts@data[ grep("Shrop", la01.pnts@data$ZONE_LABEL),]
points(la01.pnts, col = "green")
```

Central Befordshire:

```{r}
plot(la11)
plot(la11[ grep("Central B", la11$NAME), ], col = "red", add = T)
Chest <- la01.pnts[la11[ grep("Central B", la11$NAME), ],]
points(Chest)
Chest # names make sense?
Chest <- row.names(Chest)
la01.pnts[Chest,]

newChest <- SpatialPointsDataFrame(coordinates(la11[ grepl("Central B", la11$NAME), ]), match.ID= F,
                                         data = data.frame(t(c(ZONE_LABEL =1000, colSums(la01.pnts@data[Chest, 2:16])))) )
proj4string(newChest) <- proj4string(la01.pnts)

la01.pnts <- rbind.SpatialPointsDataFrame(la01.pnts, newChest )
la01.pnts@data[nrow(la01.pnts@data),1] <- as.character(la11@data$NAME[ grep("Central B", la11$NAME)])

nrow(la01.pnts)
la01.pnts <- la01.pnts[- which(row.names(la01.pnts) %in% Chest) ,]
nrow(la01.pnts)
la11@data[ grep("Central B", la11$NAME), ]
la01.pnts@data[ grep("Central B", la01.pnts@data$ZONE_LABEL),]
points(la01.pnts, col = "green")

# total zones merged
length(corRegionel) + length(Wilt) + length(Northumb) + length(Durh) + 3 + 3 + 5 + length(Chest)  
```

## Now merge the results

```{r}
la11merge <- merge(la11@data, la01.pnts@data, by.y = "ZONE_LABEL", by.x = "NAME", all.x=F, sort=F)
lam <- la11merge ; rm(la11merge)
head(lam)

# allocate cdm funding
lam$cdm <- "No"
cdt05 <- grepl("Ayl|Bright|Darl|Derby$|Exeter|Lancaster$", lam$NAME)
lam[cdt05, ] 
lam$cdm[cdt05] <- "2005" 
cdt09 <- grepl("Blackpo|Bristol|Cambridge$|Chester$|Colches|Bedf|Shrop|Southend|Sefton|Stoke|Woking$|York$", lam$NAME)
lam[cdt09, ] 
lam$cdm[cdt09] <- "2009" 

# allocate GOR
la11$GOR <- "NA"
gor <- readOGR("updata/", "GOR_st121-smaller")
la11.pnts <- SpatialPointsDataFrame(coordinates(la11), data=la11@data, match.ID=F)
proj4string(la11.pnts) <- proj4string(la11)
head(gor$ZONE_LABEL)
for(i in 1:length(gor)){
  subset <- which(la11$NAME %in% la11.pnts[gor[i,],]$NAME)
  plot(gor[i,])
  points(la11.pnts[gor[i,], ])
  la11@data$GOR[subset] <- as.character(gor$ZONE_LABEL[i])
}
```

```{r}
lam$GOR <- la11$GOR
lam$Region <- "North"
lam$Region[grepl("Midl", lam$GOR)] <- "Midlands"
lam$Region[grepl("South", lam$GOR)] <- "South"
lam$Region[grepl("Lond", lam$GOR)] <- "London"

names(lam)
lam$Allm.y <- rowSums( lam[c("Tram.y", "Train.y","Bus.y","Moto.y","Car.d.y","Car.p.y","Taxi.y","Cycle.y","Walk.y", "Other.y")])
summary(lam$Allm.y / (lam$Mfh.y + lam$Allm.y)) # around 10% work mfh 2001
lam$pCycle01 <- lam$Cycle.y / lam$Allm.y

lam$Allm.x <- rowSums(lam[c("Tram.x", "Train.x","Bus.x","Moto.x","Car.d.x","Car.p.x","Taxi.x","Cycle.x","Walk.x", "Other.x")])
summary(lam$Allm.x / (lam$Mfh.x + lam$Allm.x)) # around 5% work mfh 2011???
lam$pCycle <- lam$Cycle.x / lam$Allm.x

plot(lam$pCycle01, lam$pCycle) # make this plot nice! add regions!!!
summary(lam$pCycle)
summary(lam$pCycle01)

sd(lam$pCycle)
sd(lam$pCycle01)

sum(lam$Cycle.y) / sum(lam$Allm.y)
sum(lam$Cycle.x) / sum(lam$Allm.x)

lam$Growth <- (lam$pCycle / lam$pCycle01  - 1) * 100
lam$Abs.Growth <- (lam$pCycle - lam$pCycle01) * 100

# analysing the growth in cycling
summary(lam$Growth)
summary(lam$Abs.Growth)
length(which(abs(lam$Abs.Growth) < 1)) / nrow(lam)
cD <- which(lam$Growth < 0) # cycling Down
cU <- which(lam$Growth > 0)
length(cD)
summary(lam$Growth[cD])
summary(lam$pCycle01[cU])
summary(lam$pCycle01[cD])

length(which(lam$Growth < 0))
summary(lam$Growth[cU])
hist(lam$Growth)

library(knitr)
kable(aggregate(cbind(lam$Growth, lam$Abs.Growth), by=list(lam$Region), mean))

library(gridExtra)
p1 <- qplot(data=lam, x = Growth, fill = Region, binwidth=10) +
  xlab("Relative growth in proportion cycling (%)") + theme_classic()
p2 <- qplot(data=lam, x = Abs.Growth, fill = Region, binwidth = 0.5) +
  xlab("Absolute growth in cycling (% points)") +
#   scale_fill_discrete(guide = 'none') +
  theme_classic()
grid.arrange(p1, p2)

# what were these places like in 2001?
summary(lam$pCycle01[cD])
summary(lam$pCycle01[cU])

sum(lam$Cycle.y)
sum(lam$Cycle.x)
sum(lam$Allm.x)
sum(lam$Allm.y)

sum(lam$Cycle.x) / sum(lam$Cycle.y)
sum(lam$Allm.x) / sum(lam$Allm.y) # percent cycling

(sum(lam$Cycle.x) / sum(lam$Cycle.y)) / (sum(lam$Allm.x) / sum(lam$Allm.y))
sum(lam$Allm.y)
```

First plot of the results.
Fortification + merging.

```{r}
lamf <- la11
lamf@data <- data.frame(NAME = la11$NAME)
lamf@data <- join(lamf@data, lam)
head(lamf@data)

lam$id <- lam$NAME
laf <- fortify(lamf, region = "NAME")
head(laf)
laf <- merge(laf, lam, by="id")

gorf <- fortify(gor, region = "ZONE_LABEL")
```

Plotting.

```{r}
# map theme
theme_map <- theme_classic() + theme(axis.line = element_blank(),
                                       axis.title = element_blank(),
                                       axis.ticks = element_blank(),
                                       axis.text = element_blank())

ggplot() + 
  geom_polygon(data = laf, aes(long, lat, group, fill = Growth)) +
  scale_fill_gradient2(low="red", mid="grey", high="green", name = "Change in\ncycling (%)") +
#   scale_fill_continuous(limits=c(-40, 0, 120) low="red", high = "green") + 
  geom_path(data = gorf, aes(long, lat, group = group)) +
  theme_map
ggsave("figure/overview-la.png")
```


