A regression model to explain change in cycle commuting, 2001 - 2011
========================================================

Loading, cleaning and 'tidying' the data is often the hardest part of data analysis
problems (Wickham 2014). Fortunately this problem has already
been solved, as explained in another R script. Now we are in the privileged
position of starting from an orderly dataset containing the 'independent'
and 'explanatory' variables: $c$ is vector containing real positive numbers equal
in length to the number of zones in the study area. $c$ is the
absolute change in the proportion of commuters
recording bicycle as their primary means of travel to work between the 2001 and 2011
Census of population. There are many ($m$)
explanatory variables ($x_1 - x_m$) which seek to explain $c$ assuming linear
relationships between the 'dose' (X) and response (c):

$$ c = \sum_{m} \beta_m x_m + e$$
where e represents error in the model's prediction capabilities.

```{r, echo=FALSE}
# load("updata/lam.RData")
lam$perc.car01 <- lam$Car.d.y / lam$Allm.y # prop. driving 2001 - highly signif.
cor(lam$perc.car01, lam$Abs.Growth)
plot(lam$perc.car01, lam$Abs.Growth)
# lam$perc.car01 <- lam$Car.d.x - lam$Car.d.y # growth in driving to work # barely signif.

lam$pWAc <- lam$weightAc / ((lam$Cycle.x + lam$Cycle.y)/2)
lam$CDT <- 0
lam$CDT[ lam$cdm != "No"] <- 1
```


```{r}
names(lam)
lm1 <- lm(Growth ~ avinc + pathDPP + perc.car01 + pWAc +
            changeYMW + changeCar + CDT, data = lam, na.action="na.omit")
summary(lm1)
```

You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
plot(cars)
```

